Fluent statistical computing interfaces for biological data analysis

### 2020-07-16 Rmd/04-intron.Rmd
fix reference and delete old stuff
update bib

Although there are many flexible and powerful software tools, like BEDtools or DeepTools, for exploring and estimating coverage from common genomic data formats [@Quinlan2010-gc; @Ramirez2014-sh], it is advantageous to have tooling
that is tightly coupled to a statistical computing language such as R. This allows interoperability between other software packages for data wrangling, visualisation and modelling within the ecosystem of the language that may not
be possible with a single command line tool. There are also gains in reproducibility as analysts do not have to move between multiple software suites to explore their processed transcriptomics data.

@@

FASTQ files were aligned to the GRCz11 reference genome using subjunc with
the default parameters called from Rsubread to produce BAM files for each
sample. The coverage was then estimated directly from the set of BAM files
using **superintronic** into the long form *GRanges* representation we described above.

@@

```{r si-upset, echo=FALSE, fig.cap = "(ref:si-upset)", out.width="80%"}
knitr::include_graphics(here::here("img", "si-upset.pdf"))
```

@@

Our zebrafish workflow shows that our approach using **superintronic** 
and **plyranges** is able to uncover interesting biological signals in a
purely data-driven manner. We did not include additional information
that could have been useful when deriving our selected genes, such as
sequence motifs for U12 class of introns, or exploit the experimental
design to find differential IR like profiles. However, if that was of interest,
one could look at the overlaps, like we did in figure \@ref(fig:si-upset),
or combine our data descriptors with external estimates using **limma** 
[@Ritchie2015], like our proposed index method in @Lee2019-mf.

>>>>

Although there are many flexible and powerful software tools, like BEDtools or deepTools, for exploring and estimating coverage from common genomic data formats [@Quinlan2010-gc; @Ramirez2014-sh], it is advantageous to have tooling
that is tightly coupled to a statistical computing language such as R. This allows interoperability between other software packages for data wrangling, visualisation and modelling within the ecosystem of the language that may not
be possible with a single command line tool. There are also gains in reproducibility as analysts do not have to move between multiple software suites to explore their processed transcriptomics data.

@@

FASTQ files were aligned to the GRCz11 reference genome using subjunc with
the default parameters called from Rsubread to produce BAM files for each
sample [@Liao2013-eb; @Liao2019-qn]. The coverage was then estimated directly from the set of BAM files
using **superintronic** into the long form *GRanges* representation we described above.

@@

```{r si-upset, echo=FALSE, fig.cap = "(ref:si-upset)", out.width="80%"}
knitr::include_graphics(here::here("img", "si-olaps.png"))
```

@@

Our zebrafish workflow shows that our approach using **superintronic** 
and **plyranges** is able to uncover interesting biological signals in a
purely data-driven manner. We did not include additional information
that could have been useful when deriving our selected genes, such as
sequence motifs for U12 class of introns, or exploit the experimental
design to find differential IR like profiles. However, if that was of interest,
one could look at the overlaps, like we did in figure \@ref(fig:si-upset),
or combine our data descriptors with external estimates using **limma** 
[@Ritchie2015-mj], like our proposed index method in @Lee2019-mf.
